### 2015-04-26 evaluation.tex
various updates

\begin{table*}[t]
\small
\centering
\caption{Runtime comparison with different yield instrumentation granularities and context-switch bounds (csb).}
\label{tab:results}
\begin{tabular}{l rrr rrr rr r r}
\centering
\input{experiments/tables/scalability.tex}
\end{tabular}
\end{table*}

%\begin{table*}[t]
%\small
%\centering
%\caption{Program statistics, analysis and scalability results. Corral uses a context switch bound of 2.}
%\label{tab:races}
%\begin{tabular}{l rrr rrr rr r r}
%\centering
%\input{experiments/tables/granularity.tex}
%\end{tabular}
%\end{table*}
%
%\begin{table*}[t]
%\small
%\centering
%\caption{Analysis and scalability results. Corral uses a context switch bound of 5 and 9, respectively.}
%\label{tab:races2}
%\begin{tabular}{l rr r r rr r r}
%\centering
%\input{experiments/tables/granularity2.tex}
%\end{tabular}
%\end{table*}

\subsection{Experimental Setup}
\label{eval:setup}

All experiments were performed on a ... machine running ..., LLVM 3.5; SMACK 1.5; Z3 4.3.2; Boogie rev. 4192 and \corral rev. 534.

\subsection{Benchmarks}
\label{eval:benchmarks}

We evaluate our methodology against \sizeOfBenchmarks drivers taken from the 4.0 Linux kernel. \PDComment{Say something more about the benchmarks}

Table~\ref{tab:stats} presents statistics for all our benchmarks: lines of code (LoC); number of entry point pairs (\#Pairs); number of SMACK memory regions (\#MRs); number of racy pairs that \whoop identified (\#Racy Pairs); number of racy memory regions that \whoop reported (\#Racy MRs); and number of races that \corral discovered (\#Races Found).

\subsection{Context-Switch Granularity of Yield Instrumentation}
\label{eval:granularity}

We evaluate the scalability of our acceleration technique with respect to our benchmark sets (Table~\ref{tab:results}). All reported times are in seconds and averages of three runs. \corral was configured with a time budget of 3000 seconds per tool invocation (one for each pair of entry points); \whoop did not use a timeout.

Our experiments use the \corral context switch bounds \texttt{/k:2}, \texttt{/k:5} and \texttt{/k:9}. A higher context switch bound would result into deeper interleavings being explored and thus a larger sequentialized program. Intuitively, this means that we would see even greater speedups using information from \whoop, when exploring deeper interleavings.

@@

% We did not discover any data races with larger context switch bounds.

>>>>

\begin{table*}[t]
\small
\centering
\caption{Runtime comparison with different yield instrumentation granularities and context-switch bounds.}
\label{tab:results}
\begin{tabular}{l rrr rrr rr r r}
\centering
\input{experiments/tables/scalability.tex}
\end{tabular}
\end{table*}

We performed experiments to validate the usefulness of the \whoop approach (see Section~\ref{whoop}). We first present race-checking results from running \whoop and \corral on \sizeOfBenchmarks drivers taken from the 4.0 Linux kernel distribution\footnote{https://www.kernel.org}. We then evaluate the runtime performance and scalability of \corral and \whoop + \corral with different yield instrumentation granularities and context-switch bounds, to show that \whoop can efficiently accelerate bug-finding with \corral.

\noindent
\textbf{Experimental Setup}\xspace\xspace All experiments were performed on a ... machine running ..., LLVM 3.5; SMACK 1.5; Z3 4.3.2; Boogie rev. 4192 and \corral rev. 534.

\noindent
\textbf{Benchmarks}\xspace\xspace We evaluate our methodology against \sizeOfBenchmarks drivers taken from the 4.0 Linux kernel. To thoroughly test \whoop against the Linux driver APIs, we chose drivers from many different domains: block; char; ethernet; near field communication (nfc); universal serial bus (usb); and watchdog. We manually modeled the environment for all these drivers. This process required more than a month of work. We also focused on choosing non-trivial drivers (e.g. large number of entry points, deep nesting of entry point helper functions, complex use of locking, use of function pointers, etc).

\noindent
\textbf{Race-Checking}\xspace\xspace Table~\ref{tab:stats} presents statistics for all our benchmarks: lines of code (LoC); number of entry point pairs (\#Pairs); number of SMACK memory regions (\#MRs); number of racy pairs that \whoop identified (\#Racy Pairs); number of racy memory regions that \whoop reported (\#Racy MRs); and number of races that \corral discovered (\#Races Found) using a context-switch bound of 2.

We can see that \whoop reports many more races than \corral. This is expected, though, as \whoop over-approximates the original driver for soundness and scalability. This means that if a race exists, \whoop will report it (assuming tool and model correctness). On the other hand, \corral performs bounded checking, which means that it will report a race only under pre-defined bounds. A common problem for \corral when analyzing drivers is loop coverage. To deal with this we enable the over-approximation technique described in~\cite{lal2014powering}. This can potentially lead \corral to report false positives, but we have not seen this in practice.

We found the following data race in the generic\_nvram driver: during the \texttt{llseek} file operation, the driver is accessing the file offset (\texttt{file->f\_pos}) without using a lock. When we investigated another similar driver that uses the same API, we saw that it was protecting the same offset access with a mutex. This made us suspicious that the generic\_nvram driver might have a real race. We filed a bug report, but did not manage to get a response yet.

A common case of races that we discovered had to do with incrementing a counter without using a lock. This might happen for performance, but even benign races might lead to serious issues, especially if a compiler tries to perform optimizations assuming no races exist.

It is worthwhile to mention that we did not discover any additional data races with larger context switch bounds. This might mean that most races in these drivers require only a small amount of context-switches to manifest, or that \corral run out of resources before discovering the deeper bug.

\noindent
\textbf{Granularity of Context-Switches}\xspace\xspace Table~\ref{tab:results} presents runtime results from running \whoop, \corral and \whoop + \corral. All reported times are in seconds and averages of three runs. \corral was configured with a time budget of 7200 seconds (T.O. denotes a timeout) and a context-switch bound (csb) of 2, 5 and 9. By default, \corral instruments context-switches (i.e. \texttt{yield}) in all visible operations (denoted by \yieldall in the table). \whoop + \corral, uses two context-switch instrumentation granularities (see Section~\ref{whoop:bugfinding}): \yieldcoarse and \yieldmr.

A higher context switch bound results into deeper interleavings being explored and thus a larger sequentialized program. Intuitively, this means that we see even greater speedups using information from \whoop, when exploring deeper interleavings.

@@

### 2015-04-26 whoop.tex
various updates

Watchdog summarization uses the Houdini~\cite{flanagan2001houdini} invariant inference algorithm (available inside Boogie) to automatically compute summaries (pre- and post-conditions and loop invariants) from a pool of \emph{candidate} invariants.

@@

\subsubsection{Sound Partial-Order Reduction}
\label{bf:reduction}

By default, \corral instruments a \texttt{yield} statement (which denotes a
potential nondeterministic context-switch) before each \emph{visible operation}
(i.e., synchronization operations and shared memory accesses) of every thread.
It then leverages sequentialization to explore all possible thread interleavings
up to a pre-defined bound.

@@

\subsubsection{Race-Checking Instrumentation}
\label{bf:racechecking}

We employ a simple, but effective, encoding of data-race checks within \corral.
For example, whenever there is a write access to a shared variable, such as
\texttt{\$M.13 := 0}, we instrument the Boogie IVL code to check for write-write
races as follows:
%
\begin{boogie}
$M.13 := 0;        // original write
yield;             // allow for a context-switch
assert $M.13 == 0; // check written value
\end{boogie}
%
Likewise, for a read access to a shared variable, such as \texttt{\$p0 :=
\$M.13}, we instrument the Boogie IVL code to check for read-write races as
follows:
%
\begin{boogie}
$p0 := $M.13;        // original read
yield;               // allow for a context-switch
assert $p0 == $M.13; // check read value
\end{boogie}

>>>>

Watchdog summarization uses the Houdini~\cite{flanagan2001houdini} invariant inference algorithm (available inside Boogie) to automatically compute summaries (pre- and post-conditions and loop invariants) from a pool of \emph{candidate} invariants. The reason we need loop invariants is that we over-approximate loops using loop-cutting (also available inside Boogie).

@@

\noindent
\textbf{Sound Partial-Order Reduction}\xspace\xspace By default, \corral instruments a \texttt{yield} statement (which denotes a potential nondeterministic context-switch) before each \emph{visible operation} (i.e., synchronization operations and shared memory accesses) of every thread. It then leverages sequentialization to explore all possible thread interleavings up to a pre-defined bound.

@@

\noindent
\textbf{Race-Checking Instrumentation}\xspace\xspace We employ a simple, but effective, encoding of data-race checks within \corral.
For example, whenever there is a write access to a shared variable, such as
\texttt{\$M.13 := 0}, we instrument the Boogie IVL code to check for write-write
races as follows:
%
\begin{boogie}
$M.13 := 0;        // original write
yield;             // allow for a context-switch
assert $M.13 == 0; // check written value
\end{boogie}
%
Likewise, for a read access to a shared variable, such as \texttt{\$p0 :=
\$M.13}, we instrument the Boogie IVL code to check for read-write races as
follows:
%
\begin{boogie}
$p0 := $M.13;        // original read
yield;               // allow for a context-switch
assert $p0 == $M.13; // check read value
\end{boogie}
